Seeq: a library for inexact DNA sequence matching

### 2015-04-02 doc/document.tex
Corrected typo in manuscript.

We use a DFA representation of a dynamic programming algorithm.
We search the occurrences of the pattern $P$ of length $m$
in a text $T$ of length $n$, with up to $k$ differences.
The edit matrix $\mathbf{C} \in \mathbb{N}^{m\times n}$ is initialized
with $\mathbf{C}(0,j)=0$ and $\mathbf{C}(i,0) = i$, and then updated
column-wise with the Needleman-Wunsch recusrion. Note that this process
can be represented as $\mathbf{c}_j=f(\mathbf{c}_{j-1},T_j)$,
\textit{i.e.} the value of the $j$-th column $\mathbf{c}_{j}$ is a
deterministic transition that only depends on $\mathbf{c}_{j-1}$ and the
character $T_j$. We take advantage of this property to build a
deterministic automaton where the states are the columns of the matrix
and the transitions are dictated by the input characters.
We represent the states of the DFA by the differential encoding
of the $j$-th column, $\mathbf{s}(i)=\mathbf{C}(i+1,j)-\mathbf{C}(i,j)+1$
so that the state can be stored in a ternary trie ($\mathbf{s}(i)$
can take only 3 values).

\begin{figure}[!tpb]
\centerline{\includegraphics[scale=0.6]{Figure_1.pdf}}
\caption{Dynamic DFA construction process. In this example, the search
text is TGTGCA and the pattern is GCAC.
A) Dynamic programming matrix of the alignment. The differential code
of each column is shown inside the semicircles. B) State of the DFA at
the end of the alignment.}\label{matrix}
\end{figure}

@@

\begin{figure}[!tpb]
\centerline{\includegraphics[scale=0.5]{results.pdf}}
\caption{Benchmark between seeq, agrep and nr-grep. Each dot represents
the running time of one search with a random pattern and a random
edit distance. Note the logarithmic scale on the y axis.}
\label{fig:results}
\end{figure}

>>>>

We use a DFA representation of a dynamic programming algorithm.
We search the occurrences of the pattern $P$ of length $m$
in a text $T$ of length $n$, with up to $k$ differences.
The edit matrix $\mathbf{C} \in \mathbb{N}^{m\times n}$ is initialized
with $\mathbf{C}(0,j)=0$ and $\mathbf{C}(i,0) = i$, and then updated
column-wise with the Needleman-Wunsch recursion. Note that this process
can be represented as $\mathbf{c}_j=f(\mathbf{c}_{j-1},T_j)$,
\textit{i.e.} the value of the $j$-th column $\mathbf{c}_{j}$ is a
deterministic transition that only depends on $\mathbf{c}_{j-1}$ and the
character $T_j$. We take advantage of this property to build a
deterministic automaton where the states are the columns of the matrix
and the transitions are dictated by the input characters.
We represent the states of the DFA by the differential encoding
of the $j$-th column, $\mathbf{s}(i)=\mathbf{C}(i+1,j)-\mathbf{C}(i,j)+1$
so that the state can be stored in a ternary trie ($\mathbf{s}(i)$
can take only 3 values).

\begin{figure}[!tpb]
\centerline{\includegraphics[scale=0.6]{Figure_1.eps}}
\caption{Dynamic DFA construction process. In this example, the search
text is TGTGCA and the pattern is GCAC.
A) Dynamic programming matrix of the alignment. The differential code
of each column is shown inside the semicircles. B) State of the DFA at
the end of the alignment.}\label{matrix}
\end{figure}

@@

\begin{figure}[!tpb]
\centerline{\includegraphics[scale=0.5]{results.eps}}
\caption{Benchmark between seeq, agrep and nr-grep. Each dot represents
the running time of one search with a random pattern and a random
edit distance. Note the logarithmic scale on the y axis.}
\label{fig:results}
\end{figure}
